---
title: "A meaningful Title"
author: "Xihui Xu"
format: pdf
execute:
  echo: false
  warning: false
editor: visual
---

## Introduction

List Table/Plot Caption and formatting cross reference Appendix at the end

Useful link:

1.  https://quarto.org/docs/authoring/markdown-basics.html

2.  https://quarto.org/docs/authoring/cross-references.html

3.  https://quarto.org/docs/output-formats/html-code.html

4.  ordered list

5.  item 2

-   unordered list
    -   sub-item 1
    -   sub-item 2
        -   sub-sub-item 1

1.  ordered list
2.  item 2
    i)  sub-item 1
        A.  sub-sub-item 1

## Conclusion

@tbl-tables is a good table.

```{r}
#| echo: false
#| label: tbl-tables
#| tbl-cap: "Cars"
data(mtcars)
library(kableExtra)
kable(head(cars))
```

```{r}
#| label: fig-plot1
#| fig-cap: "Plot"
plot(mtcars$mpg, mtcars$disp, main = "meaningful title")
```

@fig-plot1 is a meaningful plot.

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

# Appendix- Include ALL your code

```{r}
#| echo: true
#| eval: false

# ALL your code here

```
